---
title: "Final Project R Code"
author: "Andrew Griffin, Madhav Mohan, Katie Taylor"
date: "6/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(Sleuth3)
library(ggplot2)
library(ggformula)
library(dplyr)
library(broom)
library(car)
library(ggthemes)
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
Prem <- read_excel("~/Math245/Prem (1).xlsx")
library(GGally)
library(kableExtra)
library(tidyverse)
```

### creating our training data set

```{r}
training_set <- filter (Prem, season< 2017)
```

### creating our validation data set

```{r}
validation_set <- filter (Prem, season>= 2017)
```

### summary stats for the prem 

```{r}
summary(Prem)
```

### scatterplot matrix to check for multicollinearity and to check certain variables if they stood out. 

```{r fig.height=50, fig.width=50}
#scatterplotMatrix( ~ points + goals + total_yellow_card + total_red_card + penalty_goals + total_shots_target + freekick_goals + in_box_goals + counter_attack + total_offside + clean_sheet + goals_conceded + saves + interception + total_tackle + total_clearance + pen_goals_conceded + total_pass + total_through_ball + total_cross + corner_taken + touches + lost_posession + champion + relegated, data = Prem, smooth = FALSE)
```



### baseline model with all variables included from dataset. 

```{r}
Model <- lm(points~ goals+total_yellow_card+total_red_card+Total_number_shots+	total_shots_target+hit_woodwork+Headed_goals+penalty_goals+freekick_goals+in_box_goals+	out_box_goals+counter_attack+total_offside+clean_sheet+goals_conceded+saves+outfielder_block	+interception+total_tackle+last_man_tackle+total_clearance+head_clearance+own_goals+	penalty_conceded+pen_goals_conceded+total_pass+total_through_ball+total_long_balls+	backward_pass+	total_cross+	corner_taken+	touches+	big_chance_missed+	clearance_off_line+	lost_posession+	penalty_save+	total_high_claim+	punches+	season+	champion+	relegated, data = training_set)
summary(Model)
```

### graphs that have been colour coded using perception to check for possible interactions. 

```{r, progress = FALSE}
#Points and goals
g1 <- ggpairs(data = Prem,mapping = aes(color = Perception , alpha = 0), columns = c("points", "goals"))
```

```{r}
#Points and clean sheet
g2 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "clean_sheet"))
```

```{r}
g3 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "goals_conceded"))
```

```{r}
g4 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "total_pass"))
```

```{r}
g5 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "total_cross"))
```

```{r}
g6 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "touches"))
```

```{r}
g7 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "champion"))
```

```{r}
g8 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "penalty_goals"))
```

```{r}
g9 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "counter_attack"))
```

```{r}
g10 <- ggpairs(data = Prem,mapping = aes(color = Perception, alpha = 0), columns = c("points", "saves"))
```

```{r}
g1
g2
```

```{r}
g3
g4
```

```{r}
g5
g6
```

```{r}
g7
g8
```

```{r}
g9
g10
```



### Model with all significant variables excluding explanatory variables, with a alpha level of 0.15,reached through backward elimination. 

```{r}
BetterModel <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded +total_cross + Perception: total_cross , data = training_set)
summary(BetterModel)
AIC(BetterModel)
better_aug <- augment(BetterModel)
gf_qq( ~ .std.resid, data = better_aug) %>%
  gf_qqline()
gf_point(.std.resid ~ .fitted, data =better_aug) %>%
  gf_hline(yintercept = 0, linetype = 2)
```

### Final model reached after adding our explanatory variables. ALong with QQ plot and residual plot to check if transformations are needed. 

```{r}
BestModel <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded + pen_goals_conceded +total_cross +  + total_red_card + total_yellow_card  + Perception: total_cross + Perception: total_red_card + Perception:total_yellow_card, data = training_set)
summary(BestModel)
knitr::kable(tidy(summary(BestModel), format = latex, getOption(2)))
AIC(BestModel)
Best_aug <- augment(BestModel)
gf_qq( ~ .std.resid, data = Best_aug, ylab = "Standardized Residuals", xlab = "N (0,1) Quantiles") %>%
  gf_qqline()
gf_point(.std.resid ~ .fitted, data =Best_aug, ylab = "Standardized Residual", xlab = "Fitted Values") %>%
  gf_hline(yintercept = 0, linetype = 2)
```

### check for multicollinearity, shows there is some but through logical conclusions this result did not make sense.  

```{r}
#GGally::ggnostic(broomify(BestModel), progress = FALSE)
#data.frame(vif(BestModel))
```

### check for outliers, there are none in training set. 

```{r}
kable(tidy(infIndexPlot(BestModel, vars = c ("Cook", "Studentized", "hat")))) 
```
### Final model compared to a model with no interactions, using anova test, shows that our final model is good.   
```{r}
NoInteractionModel <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded +total_cross +  total_red_card + total_yellow_card +  pen_goals_conceded , data = training_set)
summary(NoInteractionModel)
anova(NoInteractionModel, BestModel)
```

### table 1 
```{r}
vars <- c("in_box_goals", "out_box_goals", "clean_sheet", "goals_conceded", "total_cross", "total_red_card", "total_yellow_card", "pen_goals_conceded", "Perception")
description <- c("The goals scored inside the penalty area",
                 "The goals scored outside of the penalty area",
                 "The defending team concedes zero goals",
                 "The number of goals conceded by the team",
                 "The total number of crosses created by the team",
                 "The total number of red cards the team has received",
                 "The total number of yellow cards the team has received",
                 "The number of goals conceded through penalties",
                 "If the team is consistently perceived to be good (the big six)")

table1 <- data.frame(vars, description)
names(table1) <- c("Variable", "Description")
```

### final model based on validation set, along with QQ plot and residual plot to check if transformations are needed. 
 
```{r}
final_model <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded + pen_goals_conceded +total_cross +  + total_red_card + total_yellow_card  + Perception: total_cross + Perception: total_red_card + Perception:total_yellow_card, data = validation_set)
summary(final_model)
knitr::kable(tidy(summary(final_model), format = latex, getOption(2)))
AIC(final_model)
final_model_aug <- augment(final_model)
gf_qq( ~ .std.resid, data = final_model_aug, ylab = "Standardized Residuals", xlab = "N (0,1) Quantiles") %>%
  gf_qqline()
gf_point(.std.resid ~ .fitted, data =final_model_aug, ylab = "Standardized Residual", xlab = "Fitted Values") %>%
  gf_hline(yintercept = 0, linetype = 2)
```

### final model (validation) compared to a model with no interactions, using anova test, shows that our final model is not significant.
```{r}
NoDiscipline <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded + total_cross , data = validation_set)
anova(NoDiscipline, final_model)
NoInteraction <- lm(points~  in_box_goals + out_box_goals + clean_sheet+goals_conceded +total_cross + total_red_card + total_yellow_card + pen_goals_conceded, data = validation_set)
anova(NoInteraction, final_model)
```

### check for multicollinearity, shows there is some but through logical conclusions this did not make sense.  

```{r}
GGally::ggnostic(broomify(final_model), progress = FALSE)
data.frame(vif(final_model))
```

### check for outliers, there are none in validation set. 

```{r}
kable(tidy(infIndexPlot(final_model, vars = c ("Cook", "Studentized", "hat")))) 
```


